Explore immune cell types in Ewing sarcoma

### 2025-03-20 analyses/infercnv-consensus-cell-type/exploratory-notebooks/01_ewings-consensus-cell-types.Rmd
notebooks -> exploratory-notebooks

# Path to consensus cell types for this project
celltype_dir <- file.path(result_dir, "cell-type-consensus", "SCPCP000015")

@@

```{r paths}
# Input files
celltype_files <- list.files(
  path = celltype_dir,
  pattern = "_processed_consensus-cell-types\\.tsv\\.gz$",
  recursive = TRUE,
  full.names = TRUE
)

# Output files
```

@@

celltype_df <- celltype_files |>
  purrr::map(readr::read_tsv) |>
  purrr::list_rbind() |>
  dplyr::select(library_id, barcodes, consensus_annotation) |>
  dplyr::mutate(
    category = dplyr::case_when(
      consensus_annotation == "Unknown" ~ "Unknown",
      consensus_annotation %in% immune_celltypes ~ "immune",
      .default = "other"
    )
  )
```

@@

@@

* There are sufficient immune cells to pool together to run `inferCNV` specifying them as a normal reference
* We could also consider using only `macrophage` cells, or combined `macrophage` and all `T cell` types together, if we wanted to assess any potential effect of limit the types of immune cells feeding into `inferCNV`.

>>>>

# Path to cell-type-ewings results, which also contains consensus annotations
ewings_dir <- file.path(result_dir, "cell-type-ewings", "SCPCP000015")

@@

```{r paths}
# Input files
celltype_files <- list.files(
  path = ewings_dir,
  pattern = "_ewing-celltype-assignments\\.tsv$",
  recursive = TRUE,
  full.names = TRUE
)
```

@@

celltype_df <- celltype_files |>
  purrr::map(readr::read_tsv) |>
  purrr::list_rbind() |>
  dplyr::select(library_id, barcodes, ewing_annotation, consensus_annotation) |>
  dplyr::mutate(
    category = dplyr::case_when(
      consensus_annotation == "Unknown" ~ "Unknown",
      consensus_annotation %in% immune_celltypes ~ "immune",
      .default = "other"
    )
  )
```

@@

```{r}
celltype_df |>
  dplyr::filter(category == "immune") |>
  dplyr::count(ewing_annotation) |>
  dplyr::filter(stringr::str_detect(ewing_annotation, "tumor"))
```

Well, it appears there are some consensus immune cells which were classified as `tumor`. 
Which samples are these in?

```{r}
celltype_df |>
  dplyr::filter(
    stringr::str_detect(ewing_annotation, "tumor"),
    category == "immune"
  ) |>
  dplyr::count(library_id, ewing_annotation, consensus_annotation)
```

These results show macrophage and T cells, which are the two most prevalent categories immune cell categories.
We might want to investigate these cells further, but at least we should probably exclude them as "normal" cells for `inferCNV` since their true status is bit unclear.

@@

* There are sufficient immune cells to pool together to run `inferCNV` specifying them as a normal reference
* We could also consider using only `macrophage` cells, or combined `macrophage` and all `T cell` types together, if we wanted to assess any potential effect of limit the types of immune cells feeding into `inferCNV`. 
* There are 17 cells which are annotated as an immune consensus cell type (specifically, macrophage and mature T cell) which the `cell-type-ewings` module annotated as tumor cells.
We would probably want to exclude these cells from any normal reference.
